Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures

### 2012-02-29 ltx/paper_short_2.tex
newest draft (paper_short_2.tex) is here

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for general object tracking or tailored to a specific object type for a given application. In our implementation, we give the specific distributions for representing objects in a general manner (Section~\ref{sec:modelspecification}), which we apply to data extracted via a basic frame differencing procedure. We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

@@

A variety of methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects. These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.

@@

A great deal of research has focused on developing algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method to carry out this goal, has been used for many years, and has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach based on a Gaussian mixture model has also been used to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of the given objects in a video. The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking. This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data; our model is then applied to this noisy extracted data. Inference is performed on this model to achieve detection and tracking results.

@@

\begin{figure*}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_723.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_724.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}} 
  %../img/longimg_7_detail.pdf}} 
  \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i). The final image shows the results of frame differencing over a sequence of images (from the PETS2009/2010 dataset.}
  \label{fig:img_and_framediff}
\end{figure*}

@@

The basic set of observations extracted from video are the three-dimensional points
\begin{equation}
  \centering
  \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
\end{equation}
where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index. These data are extracted by recording the positions of pixels lying in regions that contain objects. For our implementation in experiments (Section~\ref{sec:experiments}), we desire a data extraction procedure that is as unsophisticated as possible, to verify that we can perform accurate inference on noisy data extracted using a broadly applicable procedure. Consequentially, we carry out simple frame differencing (introduced in Section~\ref{sec:introduction}) for all experiments, to find pixels at locations where there is motion. This data extraction procedure is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require data extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

We can extract additional features from videos and incorporate this new information into our model. For example, we can extract color information, and modify our model to allow for object color to be inferred during tracking. Adding this feature allows our method to better distingush between adjacent objects and more accurately track objects through occlusion. For each extracted pixel, we compute a $V$ dimensional vector that represents the color distribution in the pixel's immediate vicinity (details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}). We call the components of this vector the ``color counts'' of the pixel. The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index. 
Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.

@@

\section{Model Definition}
\label{sec:modeldefinition}

A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is applied to the extracted data. We define the generalized form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}. We also define a secondary form of this model in Section~\ref{sec:MCMC}, which is used in one of the two inference algorithms. Background information providing a brief introduction to mixture models and the Dirichlet process is contained in Appendix~\ref{sec:modelbackground}.

@@

The extracted data comprise a collection of spatiotemporal clusters, each corresponding with an object in the video. Since objects can enter and exit a scene, the number of clusters present in each time step may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps). To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture, introduced by \cite{caron_2007}. This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps. Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects in a video---to be found in an automated manner.

Similar to a typical mixture model, in this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, cluster sizes in the GPUDDPM may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$) to be
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.

We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$), to be
\begin{equation}
\begin{split}
\label{C_distro}
\text{C} ( c_{i,t} = k | &m_{1:K_{t},t}, \hspace{1mm} \alpha)\\ :=
& \begin{cases}
\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
\end{cases}
\end{split}
\end{equation}
$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$.

Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively to be
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   & \sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture. The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
\end{figure}

@@

\section{Model Specification}
\label{sec:modelspecification}

>>>>

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}). To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application. In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}). We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}). Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}). We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects. Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.

@@

A variety of methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects. These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects \cite{Okuma04aboosted,eth_biwi_00633,4036928, khan_2004, leibe2008coupled}, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002,paragios2000geodesic}.

@@

A great deal of research has focused on developing algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}. Multiple independent single-object trackers running simultaneously have been shown to be ineffective, as they will tend to coalesce and track the same object \cite{khan_2004}. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.

Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos. Blob tracking, a basic method to carry out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}. Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}. To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}. Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}. Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}. Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.

Nearly all high accuracy object detection and tracking methods are those tailored for specific object types. These methods rely on ad-hoc detection criteria that exploit knowledge about the appearance or behavior of the given objects in a video. They often make use of state-of-the-art detectors designed to locate the specified objects of interest. In contrast, the method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos. The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods. Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking. This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data. Our model of this data serves as a general framework for which we can choose a variety of object appearance distributions and inference algorithms; each choice provides a new method for unsupervised detection and tracking of arbitrary objects in videos.

@@

\begin{figure*}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_723.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_724.pdf}} \hspace{0.1mm}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame1.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame2.png}} \hspace{0.1mm}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}}
  \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i). The final image shows the results of frame differencing over a sequence of images (from the PETS2009/2010 dataset.}
  \label{fig:img_and_framediff}
\end{figure*}

@@

% The basic set of observations extracted from video are the three-dimensional points
% \begin{equation}
%   \centering
%   \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
% \end{equation}
% where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index. These data are extracted by recording the positions of pixels lying in regions that contain objects. For our implementation in experiments (Section~\ref{sec:experiments}), we desire a data extraction procedure that is as unsophisticated as possible, to verify that we can perform accurate inference on noisy data extracted using a broadly applicable procedure. Consequentially, we carry out simple frame differencing (introduced in Section~\ref{sec:introduction}) for all experiments, to find pixels at locations where there is motion. This data extraction procedure is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require data extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

We desire a data extraction procedure that yields observations of the form
\begin{equation}
\centering
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image characteristics in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.

We'd like to use an extraction procedure that is as unsophisticated as possible. Consequently, we use frame differencing. This procedure locates the pixels in image regions where there is motion relative to a previous frame. Here, each pixel corresponds to an observation $\bold{x}$. Frame differencing is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (the videos used in experiments are stationary; moving-camera videos require data extraction methods useful for non-stationary video \cite{chien2002efficient, zhang2007moving}). Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).

We also extract features $\bold{x}^{c}$ that capture image information in the vicinity of each extracted pixel. Examples of possible features include color distributions, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations. In principle, we can extract any image features that may be used to characterize the appearance of objects. In our implementation, we choose to extract only color information in the vicinity of each pixel. Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distingush between adjacent objects and track objects through occlusion. To add this information, we let $\bold{x}^{c}$ represent a $V$ dimensional discrete distribution over some aspect of color (such as the hue) in the pixel's immediate vicinity. Details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}. We refer to the components of this vector as the ``color counts'' of the pixel.

% old discussion on incorporating color features

% The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
% \begin{equation}
% \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
% \end{equation}
% where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index.
% Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.

% old (long) data extraction - parts grabbed for new version above
% ----------------------------------------------------------------

@@

\section{Model Framework}
\label{sec:modeldefinition}

We use a type of dependent Dirichlet process mixture (DDPM) model known as the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM). We define a general form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}. We also define a secondary form of this model in Section~\ref{sec:MCMC}, which is used in one of the two inference algorithms. We provide a brief introduction to mixture models and the Dirichlet process in Appendix~\ref{sec:modelbackground}.

Dirichlet process mixture (DPM) models (Section \ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric models. These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis. Data extracted from videos comprises spatiotemporal clusters, each corresponding with a distinct object. Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data. 
%In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.

Since objects can enter and exit a scene, the number of clusters present throughout the video may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps). To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture, introduced by \cite{caron_2007}. This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps. Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be found in an automated manner.

@@

In this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
%In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
The size (i.e. number of assigned observations) of each cluster in the GPUDDPM may decrease over time as observations become ``unassigned''. We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$),
\begin{equation}
\begin{split}
\label{D_distro}
\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.

We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$),
\begin{equation}
\begin{split}
\label{C_distro}
\text{C} ( c_{i,t} = k | &m_{1:K_{t},t}, \hspace{1mm} \alpha)\\ :=
& \begin{cases}
\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
\end{cases}
\end{split}
\end{equation}
$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$.

Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively as
\begin{align}
\label{gpuddpm_def}
\begin{split}
m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
\theta_{k, t} | \theta_{k, t-1}   & \sim
\begin{cases}
  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
\end{cases} \\
c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
\end{split}
\end{align}
$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}. The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
\begin{figure}[h]
        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture. The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
\end{figure}

@@

\section{Model Specifics}
\label{sec:modelspecification}
